Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-08-15 paper_jasa.tex
updated raference and abstract
tweaked the abstract

\graphicspath{{images/}}
\usepackage{color}
\newcommand{\blue}[1]{{\color{blue} #1}} %MM
\newcommand{\red}[1]{{\color{red} #1}} 
\newcommand{\green}[1]{{\color{green} #1}} %DC

\definecolor{orange}{rgb}{1,0.5,0}
\newcommand{\hh}[1]{{\color{orange} #1}} %HH

@@

% Article top matter
\title{Validation of Visual Statistical Inference, Applied to Linear Models}
\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
\thanks{Mahbubul Majumder is Graduate Student (e-mail: mahbub72@gmail.com) , Heike Hofmann is Associate  Professor and Dianne Cook is Professor, Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210. This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
\date{\vspace{-.7in}}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when \citet{buja:2009} conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.

>>>>

\graphicspath{{images/}}

@@

% Article top matter
\title{Validation of Visual Statistical Inference, Applied to Linear Models}
\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210. This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
\date{\vspace{-.7in}}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract}  
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. A new method, called the lineup protocol, enables determining the statistical significance of visual findings, which bridges a big gulf between exploratory and inferential statistics.  In this paper inferential methods for statistical graphics is developed further by refining the terminology of visual inference, and framing the lineup protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists. This framework is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides comparable results to the conventional tests, and expectedly out-performs them when data is contaminated, a scenario where assumptions required for performing a conventional are violated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult of tasks.
